\chapter{Data-driven Networking}\label{chap:ddn}
%Once we have any computer network as described above, we have a good basis for not only allowing two machines to communicate with one another, but also to do so \emph{efficiently}.
%That isn't to say that there is no room to improve on how networks and operating systems provide this functionality.
%In amongst the design of ?? \cref{chap:nets} are ?? tunable aspects, algorithms in protocol to improve perf... ????
Many aspects of modern networks, such as \glspl{acr:cca}, thresholds for differentiating services, and flow classification rely heavily upon hand-tuned heuristics.
As in \cref{sec:in-network-compute-use-cases}, there is still vast scope to improve on communication latency and throughput, or to avoid and work around deleterious traffic patterns (such as incast communication).
As a result, research into \glspl{acr:cca}, network designs, and routing procedures is very much ongoing.
Crucially, as these operations lie at the core of network operation their solutions tend towards extremely efficient heuristic methods; they must be evaluated per-packet or react as quickly as possible to state change.
Unlike the approaches empowered by the programmability explored through \cref{chap:nets}, the norm is that most deployment environments have no access to such tools, and thus have reduced scope for co-design or measuring network properties that would make their design simpler.
Designing new methods for network optimisation then requires deep insight into any problem, its edge cases, and the hardware \& performance characteristics of the target environment.\sidenote{This affects how we design such heuristics even on commodity hosts---for instance, kernel-space \glspl{acr:cca} (i.e., as part of the TCP stack) are unable to use floating point arithmetic.}

Suppose that, as network administrators or protocol designers, we have access to a reasonable amount of information about the machines, network segments or \glspl{acr:as} under our control---measurements, observations, and statistics taken at run-time, from simulation, or by modelling.
A natural question to ask, then, is whether we can use this data to enhance and improve the operation and use of our network automatically.
Thinking further still, we might wonder whether we can outperform the general (yet useful) heuristics which are widely deployed and researched, allowing us to tailor network behaviour according to its environment and traffic patterns.
These questions are the founding principles of \glsxtrfull{acr:ddn}\sidenote{Alternatively titled \emph{self-driving networks}.}, a recent field of research focussed on the automatic control and optimisation of network systems, which has sprung forth due to recent advances in \gls{acr:ml} and \gls{acr:rl}~\parencite{DBLP:conf/anrw/FeamsterR18,DBLP:journals/pieee/KellererKBBR019}.

The ideas and goals of automated network control have always existed and evolved in one form or another, particularly as computational inference and learning have grown more powerful.
Primarily, these ideas have propagated in their early forms via position papers offering a `vision of things to come'.
This was first famously formalised as the \emph{knowledge plane}~\parencite{DBLP:conf/sigcomm/ClarkPRW03}, in contrast to the \emph{data} and \emph{control} planes.
This proposal captures not only the above concepts of automation as a means for network control, but also for collaborative or commercial sharing of information between end-hosts, transit \glspl{acr:as}, and organisations to build up a global picture of the needs of the network.
In truth, over the past \num{19} years we have moved no closer to such a unified substrate, though automated inference based on the data we \emph{do} have is richly researched.
A later attempt to combine this with \gls{acr:sdn} as \emph{Knowledge}-defined networking~\parencite{DBLP:journals/corr/MestresRCBASMMB16} takes key steps in clarifying the field, through concrete problems and promising \gls{acr:ml} developments, but effectively curtails the scale of knowledge sharing.
\gls{acr:ddn} itself is named and defined by \Textcite{DBLP:conf/comsnets/JiangSSZ17}, who again expand the scope for optimisation beyond network control to include \emph{end-points}; towards application and transport layer optimisation for hosts and servers, as well as control of the underlying fabric.

%?? Para on what is all offered here? or interpose below with `why'?

Starting out with the aim of emphasising and motivating the value of \gls{acr:ddn}, I discuss and introduce some of the recent developments and applications of \gls{acr:ml} and \gls{acr:rl} techniques in computer networking (\cref{sec:use-cases}), before then moving onto to explain the `building blocks' underlying these approaches.
Specifically, I introduce relevant function approximations (\cref{sec:function-approximation}), techniques to learn these representations including \gls{acr:rl} (\cref{sec:learning-an-approximation}), and representations for different target devices (\cref{sec:numerical-representations-for-embedded-ml}).\sidenote{There are other intersections between networked systems and data-driven techniques which I omit here but deserve a cursory mention: advancements in \gls{acr:fl} and distributed \gls{acr:ml} training, which show how computer networks can aid the wider \gls{acr:ml} field.}
\Cref{sec:ddn-challenges} discusses additional challenges inherent to \gls{acr:ddn}, while \cref{sec:ddn-security} then presents an overview of the security perspective surrounding current \gls{acr:ml} and \gls{acr:ddn} approaches.
Although this context and its challenges are rapidly evolving, an understanding of security issues is key to offering a complete picture of the viability of \gls{acr:ddn} and the care which must be taken in its research.
Sadly, full examination and further development lies beyond the scope of this thesis---it is, in fact, a thesis-worthy topic in its own right~\parencite{papernot-thesis}.

\section{Use cases}\label{sec:use-cases}%
\input{chapters/p1-introduction/ch3-ddn/use-cases.tex}

\section{Function approximation}\label{sec:function-approximation}
\input{chapters/p1-introduction/ch3-ddn/fn-approx.tex}

\section{Learning an approximation}\label{sec:learning-an-approximation}%
\input{chapters/p1-introduction/ch3-ddn/learning-approx.tex}

%\section{Collaborative training}\label{sec:ddn-collaborative-training}
%\input{chapters/p1-introduction/ch3-ddn/distributed-ml.tex}

\section{Numerical representations for embedded ML}\label{sec:numerical-representations-for-embedded-ml}
\input{chapters/p1-introduction/ch3-ddn/representations.tex}

\section{Challenges}\label{sec:ddn-challenges}
\input{chapters/p1-introduction/ch3-ddn/challenges.tex}

\section{Security}\label{sec:ddn-security}
\input{chapters/p1-introduction/ch3-ddn/security.tex}

\section{Summary}
I have introduced and discussed a wide variety of problems in the networking domain where \gls{acr:ml} and \gls{acr:rl} techniques are able to provide tangible improvements in performance and resilience of modern networks.
We have also covered the basics of \gls{acr:ml} techniques and function approximations, as well as the intuition and \gls{acr:mdp} formalism underpinning \gls{acr:rl} algorithms.
Finally, we have examined modern \gls{acr:rl} algorithms, concessions and design decisions required to bring \gls{acr:rl} and \gls{acr:ml} to network devices, and open challenges pertinent to \gls{acr:ddn}.

Unfortunately, the security context around \gls{acr:ddn} remains an uncomfortably open question.
While not all the attacks presented can \emph{immediately} generalise to networks or other problem-space representations, the absence of research into specific attacks in this field does not inspire confidence.
The interaction model of networks \emph{does} present a useful level of indirection, such that model extraction and similar approaches are harder to perform.
For instance, while exact traffic routing or queueing decisions should enjoy some degree of isolation from the actual effects they cause on a monitored flow---there are many potential hypotheses behind a decrease in maximum flow performance.
Efficacy aside, this casts significant doubt on whether we can (safely) elevate \gls{acr:rl} or other online approaches from an interesting research question towards real-world deployment.
