\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
%\def\theHchapter{2}

First of all, I'd like to thank my wife and best friend Vanessa, for her tireless support and persistence in hearing my a) writing woes, b) angry tirades about software in general, and c) despair over experiment fixes running long into the night---most often in the run up to conference deadlines.
I couldn't have made it this far without you.
I also owe my family an inestimable debt for their their lifelong help and guidance: to my parents Elaine and Ronnie for always being there and always looking our for me (including making a mid-pandemic move-in possible), my brother Rhys for being there to bounce ideas off of (and hold many technically-involved emulator discussions), and my sister Brooklyn for her constant support.
You've all always believed in me and in what I could achieve, and I can't ever pay that back. 
Lastly, Charlie\sidenote{Unabbreviated: Beloved Sir Charles of House Cattington, Esq., first of his name.}, the best cat, for providing many loud, quizzical meows and for bestowing upon me the highest honour as Chief Feeder.

From Netlab, I'd like to thank Prof.\ Dimitrios Pezaros for his guidance, trust, and advice over the years, as well as the rest of my supervisory team---Simon Rogers and Angelos Marnerides---for their input.
To the other Netlab PhD students in particular, Mircea Iordache-\c{S}ic\u{a}, Stefanos Sagkriotis, and Haruna Umar Adoga: our daily meetings have been an essential wellspring of emotional support for me all through the pandemic.
It's been great to look out for one another, in terms of morale and to keep engaged and challenged on what we're each working on.
Not to mention the many proof-readings of various lumps of text, lovingly provided free-of-charge!

For many a fine lunch at \emph{The Wee Curry Shop} and vibrant discussion (albeit on algorithmics and academic politics!), I have to thank Patrick Prosser, Ciaran McCreesh, James Trimble, Craig Reilly, Blair Archibald, and Ruth Hoffmann.
To Patrick in particular, thank you for showing me that research was somewhere I could be and succeed in, and for helping put me down this path.

I'd like to thank all of the co-residents of F101 through the years; Yousef Alhaizaey, Dhahi Alshammari, Dejice Jacob, Thomas Koehler, Lito Michala, Sangkyu Park, Adrian Ramsingh, and Cris Urlea.
The discussions, feedback on ideas and GLASS talks, technical help, and an occasionally borrowed poster tube have all been greatly appreciated.

I've met and known so many amazing people through the department via lunches, cryptic crosswords, puzzles, and Friday coffees that are but a distant memory now.
A lot of us started around the same time---or at least met constantly through boardgames, D\&D, and mutual friends: Benjamin Bumpus, Marco Cook, Frances Cooper, Patrizia Di Campli San Vito, Natascha Harth, Ellen \& William Kavanagh, Charlie Rutherford, Jess Ryan, G\"{o}zel Shakeri, Lovisa Sundin, and Tom Wallis.
Studying was that much livelier for having you around, and you made Lilybank Gardens somewhere worth going for 5 years.
To Andrew Doctor, Sami Kelly, Shaun Lithgow, Callum Milne, Jen Patrick, and Simon Tait: thank you for being lifelong friends (and keeping me sane and socialised!).
There are so many other folks I've met through the department, Waterdeep, tea society, and teaching that have been great all around: Vivian Band, Simon Fowler, Katie Heeps, George Holt, Shaun Macdonald, David Maxwell, Michael McKay, Murray McKinstray, Flora McNulty, Stephen McQuistin, Alex Pancheva, Iulia Paun, Will Pettersson, Rebecca Rae, Jenny Savage, Michel Steuwer, Ritchie Walker, and Mihail Yanev.

To Derek `Del' Hamilton, Douglas MacFarlane, Stewart MacNeill, Robert Nugent and the SoCS support staff, thanks for putting up with our unceasing demands around keeping G131 setup with whatever shiny toys we happened to acquire.
To Helen Border and Gail Reat, thanks for keeping teaching and demonstrating as such a tightly run ship.

I met some fantastic people at ESnet, whom I must thank for an incredible 3 month stint in Berkeley, exciting work, and great hills to run on.
Crucially, you helped reshape my research direction and introduced me to the joys of programmable dataplane hardware.
Richard Cziv\'{a}, Chin Guok, Yatish Kumar, Bruce Mah, Inder Monga; I hope to work with you again one day soon.

Between various online pastimes---namely, FFXIV and all manner of tabletop shenanigans---I've known and sometimes drifted away from a great number of folk who've offered distant support in their own way.
Thank you Bernhard, Cam, Cameron, Carlos, Charles, Fred, Ian, James, Katie, Kyle, Lily, Mal, Olivia, Sasha, Sphenio, and Ves.
For routinely keeping me up 'til 2AM---and later---but making it all worthwhile in exchange.

Additional thanks go to Dimitrios Pezaros, Jeremy Singer, and Haruna Umar Adoga for proof-reading part or all of this thesis.

And at last, a begrudging `thanks' to my anonymous reviewers over the years, who have been ever as much my mentors as they have my torturers and gaolers.
For the \emph{constructive} advice, I can only thank you.
