% !TeX document-id = {29d93272-0c70-47c0-a2c9-f98f7fffc7db}
% !TeX program = xelatex
% !BIB program = biber

% \documentclass[twoside,hidelink]{glasgowthesis}
\documentclass[twoside,hidelinks]{glasgowthesis}
%\documentclass[nogutter,hidelinks]{glasgowthesis}
%\documentclass[11pt,paper=b5,footinclude,headinclude]{scrbook}
%\usepackage{classicthesis}

% ^^ draft shows overfilled boxes.
% ^^ replace "oneside" with "twoside" to set the gutter correctly for
%    two-sided printing.
% ^^ add "nogutter" option for digital copy (without binding offsets),
%    if printed copy is twoside, use [twoside,nogutter] for digital copy.

\input{"preamble/packages.tex"}
\input{"preamble/bibresources.tex"}
\input{"preamble/fonts.tex"}
\input{"preamble/haxx.tex"}
\input{"preamble/math-terms.tex"}
\input{"preamble/header-cfg.tex"}
\input{"preamble/uofg-colours.tex"}

% Need to call 'makeglossaries'...
\setabbreviationstyle[acronym]{long-short}
\makeglossaries
\input{"glossary.tex"}
\input{"acronyms.tex"}
\glsenableentrycount

\newcommand{\mytitle}{Programmable, Data-driven Networks for the Masses}
\newcommand{\myname}{Kyle Andrew Simpson}

\hypersetup{
	colorlinks,
	%draft,
	citecolor=kthesis-cite,
	filecolor=black,
	linkcolor=kthesis-internal,
	urlcolor=kthesis-url,
	linktoc=all,
	pdftitle={\mytitle{}},
	pdfauthor={\myname{}}
}

% =============================================================================
\begin{document}

\pagestyle{empty}

\pagenumbering{gobble}

\title{\mytitle{}}
\author{\myname{}}
\date{December 2021}

\maketitle

\cleardoublepage

%\fontfamily{alegreya}\selectfont
%\Alegreya

\newgeometry{textwidth=11.75cm, textheight=24.3cm}
\topmargin -0.74cm
\fancyhfoffset[E,O]{5pt}

\pagenumbering{roman}

\input{"chapters/frontmatter/abstract.tex"} 

\newpage
\vspace*{1.75in}
\begin{flushright}
	\emph{
		\begin{quotation}
			\noindent
			I'm astounded whenever I finish something.
			Astounded and distressed.
			My perfectionist instinct should inhibit me from finishing; it should inhibit me from even beginning.
			But I get distracted and start doing something.
			What I achieve is not the product of an act of will but of my will's surrender.
			I begin because I don't have the strength to think; I finish because I don't have the courage to quit.
			This book is my cowardice.
		\end{quotation}
		---Fernando Pessoa, \emph{The Book of Disquiet} (pp. 156)}
\end{flushright}
\input{"chapters/frontmatter/acks.tex"}

\newpage
\input{"chapters/frontmatter/original.tex"} 

\newpage
\input{"chapters/frontmatter/typesetting.tex"}

\cleardoublepage


%\pagenumbering{roman}
%\setcounter{page}{1}

% -----------------------------------------------------------------------------
%\newpage

%\vspace*{1.75in}
%\begin{center} {\bf Education Use Consent}\end{center}
%\noindent I hereby give my permission for this project to be shown to other
%University of Glasgow students and to be distributed in an electronic
%format. {\bf Please note that you are under no obligation to sign this
%  declaration, but doing so would help future students. }

%\begin{description}
%\item [Your Name]\ \\ \ \\ \ \\
%\HRule
%\end{description}
%\addcontentsline{toc}{chapter}{Education Use Consent}

% -----------------------------------------------------------------------------
%\newpage

% -----------------------------------------------------------------------------

% -----------------------------------------------------------------------------
%\newpage
%\thispagestyle{empty}
\microtypesetup{protrusion=false}
\pdfbookmark[chapter]{\contentsname}{toc}
{\hypersetup{hidelinks}\sffamily\tableofcontents}
\listoftables
\listoffigures
\microtypesetup{protrusion=true}

%\printglossary
\setlist[description]{leftmargin=!, labelwidth=4em}
\printglossary[type=\acronymtype]
\setlist[description]{style=standard}
\setlist[description]{font={\itshape\bfseries\sffamily}}

% -----------------------------------------------------------------------------
\cleardoublepage
% Reset page numbering
\pagestyle{fancy}
\pagenumbering{arabic}
\setcounter{page}{1}
% Number chapter boomarks
\bookmarksetup{numbered}


% -----------------------------------------------------------------------------
% Introduction and research problem
%\part{Background}
\input{"chapters/p1-introduction/intro.tex"}

% -----

%\part{Background}
\input{"chapters/p1-introduction/lit-networks.tex"}
\input{"chapters/p1-introduction/lit-pdp.tex"}
\input{"chapters/p1-introduction/lit-ddn.tex"}

% -----

%\part{Improving Deployability}

\chapter{DDoS Prevention by Multi-Agent Reinforcement Learning}\label{chap:ddos-rl}
?? Something to elaborate on versus paper: design of Ryu controller?

\section{Threat Model}
\section{Agent Design}
\subsection{Weaknesses in Prior Art}
\subsection{Feature Spaces}
\subsection{Reward Functions}
\subsection{Action Spaces}
\subsection{Systems and Threat Considerations}
\section{System Design}
\section{Modified Learning Algorithms}
\section{Methodology}
\subsection{Traffic Modelling}
\subsection{Topologies and Testing Environments}
\section{Evaluation}
\subsection{Congestion-aware Traffic}
\subsection{Congestion-unaware Traffic}
\subsection{Increased Attack Volume}
\subsection{Computational Cost}

\chapter{In-network Reinforcement Learning}\label{chap:in-net-rl}

?? Note worth having here (and the paper itself): past `parallel Sarsa'~\parencite{DBLP:conf/aamas/GroundsK07} means `learning from parallel agents' traces': OPaL enables this too to some extent!

\section{Analysing State Slippage}

\section{Design}

\subsection{Functional Requirements}

\subsection{SmartNIC Targeted Designs}

\subsection{Other Proposals}

\section{Methodology}

\subsection{Topologies and Testing Environments}

\section{Evaluation}

\subsection{Quantisation}

\subsection{Effects on Background Traffic}

\subsection{Scaling with Device Resources}

\subsection{State Slippage}

\subsection{Training Time}

\subsection{Accuracy in MARL anti-DDoS}


% -----

%\part{Enabling New Use Cases}

\chapter{Scalable Flow Classification}\label{sec:ch-seidr}
\section{Motivation}
\section{Sei\dh{}r Histograms}
\subsection{Algorithm}
\subsection{Packet Generation on PSA}
\section{Use case: TCP CCA Detection}
\subsection{Observable Differences}
\subsection{Investigating the BBR Algorithm}
\section{Methodology}
\subsection{Testing Environments}
\subsection{Data Collection/Generation}
\subsection{ML Model Architecture}
\section{Evaluation}
\subsection{Device Memory Costs}
\subsection{Bandwidth Costs}
\subsection{CCA Detection Accuracy and Costs}

%\chapter{Dynamic Fault Detection}

%\part{Conclusions}

\chapter{Conclusion}

% -----------------------------------------------------------------------------

\part{Appendices}
\appendix

\input{"chapters/appendices/appendix-1.tex"}

% -----------------------------------------------------------------------------
\newpage
%\makeatletter
%\let\ps@plain=\ps@fancy
%\makeatother


% Bibliography
\addcontentsline{toc}{part}{References}
\renewcommand*{\mkbibnamefamily}[1]{\textsc{#1}}
\renewcommand*{\mkbibnameprefix}[1]{\textsc{#1}}
\printbibliography[title={References}]

\end{document}

